Taxi Trip Density Estimation in NYC\thanks{2016 Applied Statistics Qualifying Exam report.}

### 2016-05-25 paper.tex
Adjusting dpi.
Adding oos fit evaluation.
Adding oos model diagnostics.

\begin{figure}[htb] \centering
  \includegraphics[width=0.8\linewidth]{./include/time.png}
  \caption{Taxi pickups over time with proposed time categories.}
  \label{fig:time}
\end{figure}

@@

\begin{figure}[htbp] \centering
  \begin{subfigure}[t]{.49\linewidth}
    \includegraphics[width=\linewidth]{./include/gmm_cat_morn.png}
    \caption{7am-9am} \label{fig:dist:morning}
  \end{subfigure}
  \begin{subfigure}[t]{.49\linewidth}
    \includegraphics[width=\linewidth]{./include/gmm_cat_night.png}
    \caption{10pm-2am} \label{fig:dist:night}
  \end{subfigure}
  \begin{subfigure}[t]{.49\linewidth}
    \includegraphics[width=\linewidth]{./include/gmm_cat_morn2.png}
    \caption{7am-9am} \label{fig:dist:morning2}
  \end{subfigure}
  \begin{subfigure}[t]{.49\linewidth}
    \includegraphics[width=\linewidth]{./include/gmm_cat_night2.png}
    \caption{10pm-2am} \label{fig:dist:night2}
  \end{subfigure}
  \caption{Some caption.}
  \label{fig:dist}
\end{figure}
% Morning more: Southwest Brooklyn, North Queens, Bronx, North Manhattan
% Night more: Northern Brooklyn (Williamsburg)

>>>>

\begin{figure}[htb] \centering
  \includegraphics[width=0.8\linewidth]{./include/time.pdf}
  \caption{Taxi pickups over time with proposed time categories.}
  \label{fig:time}
\end{figure}

@@

\begin{figure}[htbp] \centering
  \begin{subfigure}[t]{.49\linewidth}
    \includegraphics[width=\linewidth]{./include/gmm_cat_morn.png}
    \caption{7am-9am} \label{fig:dist:morning}
  \end{subfigure}
  \begin{subfigure}[t]{.49\linewidth}
    \includegraphics[width=\linewidth]{./include/gmm_cat_night.png}
    \caption{10pm-2am} \label{fig:dist:night}
  \end{subfigure}
  \begin{subfigure}[t]{.49\linewidth}
    \includegraphics[width=\linewidth]{./include/gmm_cat_morn2.png}
    \caption{7am-9am} \label{fig:dist:morning2}
  \end{subfigure}
  \begin{subfigure}[t]{.49\linewidth}
    \includegraphics[width=\linewidth]{./include/gmm_cat_night2.png}
    \caption{10pm-2am} \label{fig:dist:night2}
  \end{subfigure}
  \caption{Some caption.}
  \label{fig:dist}
\end{figure}

\subsection{Evaluating Fit}
In order to evaluate the fit of these models, I use a test dataset consisting of 426,194 observations from December 15, 2015. I consider three models. The first is the standard GMM model described in Equation~\ref{eq:gmm}. The second, which I call subset-GMM (ssGMM), is the same GMM model estimated separately on each of the six time category subsets. And the third is the GMM model with time-varying mixing proportions (tvGMM) as described in Equation~\ref{eq:gmm_cat}.

Table~\ref{tab:logliks} presents test data diagnostics for these three models. For each model, I calculate the log-likelihood contribution for each time period, as well as the total log-likelihood and the Bayesian information criterion (BIC) for each model.\footnote{BIC is calculated as $p \log n - 2 \log L$, where $p$ is the number of estimated parameters, $n$ is the number of observations, and $L$ is the likelihood of the model. Lower BICs indicate better models.}

\begin{table}[htb] \centering
\begin{tabular}{rrrr}
  \toprule
   & \multicolumn{1}{c}{GMM} & \multicolumn{1}{c}{ssGMM} & \multicolumn{1}{c}{tvGMM} \\
  \midrule
  02:00-07:00 &  125126 &  \textbf{128015} &  127873 \\
  07:00-09:00 &  235173 &  236006 &  \textbf{237738} \\
  09:00-16:00 &  773825 &  769287 &  \textbf{780562} \\
  16:00-18:00 &  211914 &  210956 &  \textbf{214326} \\
  18:00-22:00 &  601819 &  600960 &  \textbf{606997} \\
  22:00-02:00 &  292844 &  296174 &  \textbf{298605} \\
  \midrule
  Total log-lik &  2240702 &  2241399 &  \textbf{2266101} \\
  BIC           & -4479083 & -4468876 & \textbf{-4528002} \\
  \bottomrule
\end{tabular}
\caption{Out-of-sample log-likelihoods.}
\label{tab:logliks}
\end{table}

We can see that overall tvGMM has significantly better out-of-sample fit than either GMM or ssGMM, both in terms of log-likelihood and BIC. In fact, tvGMM even outperforms ssGMM on all but one of the subsets considered.
